---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
```

# **Global Perspective On Risk Factors Associated with Adolescent Birth Rate**

By Hannah Miller and Shyhiem Lawrence

*MSUM Quantitative Biology Students*

**Abstract:**

Adolescent birth rate is the amount of births per 1000 females who give birth between the ages 15 to 19. There are many possible factors that could contribute to a high rate of adolescent birth. A high adolescent birth rate could result in mortality for both mothers and infants while this is possible factors such as truncated education, lower future family income and larger family size are also effects of adolescent birth. If we can understand the relationships between risk factors and adolescent birth rate , then we can make recommendations about how to reduce adolescent birth. But, we don't know enough about those relationships.The focus of this project is to explore the characteristics of countries that experienced high adolescent birth rate and to see how the frequency of those characteristics varied geographically. To do that, we will use data from the World Health Organization - Global Health Organization which contains information about the countries and their birth rates throughout the years 2000 to 2019. Investigation obtained from this data will be used to provide adequate information in high risk areas for adolescent birth rate.

**Introduction:**

The adolescent birth rate measures the annual number of births to women 15 to 19 years of age per 1,000 women in that age group. Adolescent birth rate has been a striking obstacle around the world, some depriving women and babies gain a better experience of life.  Social determinants of health, such as low education and low income levels of a teen's family, may contribute to the high teen birth rates. Teens in certain areas are at higher risk of teen pregnancy and birth than other areas. Not only can this pose as a risk for better experience, but it can also increase a risk of mortality in babies and women when an adolescent birth occurs.

  

**Methods:**

![](Screenshot%20(1).png)

```{r} # Histogram}
abr %>% 
  filter(year == 2018) %>% 
  ggplot() +
  geom_col(mapping = aes(x = country, y = abr), fill = "royalblue4")  +
  coord_flip()

```

```{r} # World Map}
ne_countries(scale = "medium", returnclass = "sf") %>%
  filter(sov_a3 != "ATA") %>% 
  left_join(filter(abr, year == 2018), by = c("admin" = "country")) %>% 
  ggplot() +
  geom_sf(aes(fill = abr)) +
  scale_fill_viridis_b() +
  coord_sf(crs = st_crs("ESRI:54030")) +  # Robinson
  theme_void()

```

**Results:**

<img src="https://lh6.googleusercontent.com/GymQhJKCdND8JCJxrevu961VtzgyMs-QdgQjkX7LqEtUYxeBXKMpZun_opPcOhoZeX9BGpU7WO_SDf8-G_T7wX5So7wJtyieh_YVMJOWM331jZhUF4BVsZ3psaXrlFMQoRqLFyeaNLnV" width="516"/>

**Figure 1.** A Histogram representing the Adolescent Birth Rate (abr) in the year of 2018 among different countries around the world.

![](https://lh6.googleusercontent.com/7zW4jkyjxWCmIAhlCqD2XM1VYvjSnm0FjZum4T_79ZTOckM3RcPdHsKdYWEf-QJkB-v8QhNUAF3fPK3A_9_el6bpQx8qeAKEwuLxnk7gU-95ceh1vHiKHW2hRdgFy8Tol1fP2Mfijx3Q)
**Figure 2.** A World Map showing the high and low birth rates among adolescents aged 15 to 19. Yellow representing the highest rates, while purple represents the lowest rates.

**Discussion:**

Adolescent birth rate is a striking factor that is inconvenient for both mothers and babies such as not much exposure to education.Our findings demonstrated higher adolescent birth rates in low to medium income countries in comparison to other countries. This statistics provides the ratio that one out of four females give birth before reaching the age of nineteen.  (Carvajal et al. 2020) In comparison to the data there are both highs and lows that differ throughout the year, this could be a result of not basely from developing country or not but income inequality, a country that is developed does not necessarily means that everyone is well paid hence a similarity in small income will vary between a developed and developing countries (Nagata 2016). We cannot exclude the possibility that a lack of education on the specific topic could be the leading cause and fluctuations among the years, with the inclusion of socioeconomic factors as the environment a person may liveTo combat these issues providing employment opportunities for women and improved education, urbanization, increased access to contraception and a delay in the age of marriage and childbearing could help to decrease adolescent birth rate globally.   

**References:**

"Adolescent Birth Rate (per 1000 Women Aged 15-19 Years)." *World Health Organization*, THE GLOBAL HEALTH OBSERVATOR, <https://www.who.int/data/gho/data/indicators/indicator-details/GHO/adolescent-birth-rate-(per-1000-women-aged-15-19-years).>

"Adolescent Pregnancy." *World Health Organization*, World Health Organization, <https://www.who.int/news-room/fact-sheets/detail/adolescent-pregnancy.>

Carvajal, L., Wilson, E., Requejo, J. H., Newby, H., Eriksson, C. D. C., Liang, M., Dennis, M., Gohar, F., Kapeu, A. S., Idele, P., & Amouzou, A. (2020). Basic Maternal Health Care Coverage Among Adolescents in 22 Sub-Saharan African Countries with High Adolescent Birth Rate. *Journal of global health*, *10*(2). <https://doi.org/10.7189/jogh.10.021401>

Nagata, Jason M. "Global Health Priorities and the Adolescent Birth Rate." *Journal of Adolescent Health*, vol. 60, no. 2, 16 Nov. 2016, pp. 131–132.

**Acknowledgements:**

Chris Merkord

## 
